---
title: "Projet IPS"
author: "Mohamed Sneiba HAMOUD"
date: "10/2/2018"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## R Markdown

In this project, we will compute the solutions of the $\color{blue}{\text{1D quantum harmonic oscillator}}$  and we will check some of its properties. 

#### Quantum Harmonic Oscillator
##### Schrodinger equation :
                  
$$\hat{H}(z)\Psi(z) = E_n\Psi_n(z)$$

##### Resulting 1D-HO Schrodinger equation

$$(\frac {\hat{p}^2(z)}{2m} + \frac {1}{2}m\omega^2\hat{z}^2)\Psi_n = E_n\Psi_n$$

##### Solutions:
The analytic solutions is :
$$ \Psi_n(z) = \frac {1}{\sqrt{2^nn!}} (\frac {m\omega}{\pi\hbar})^\frac {1}{4} e^{-\frac {m\omega z^2}{2\hbar} }H_n(\sqrt{\frac {m\omega}{\hbar}}.z)$$

To simplify we will take the constants $m$, $\hbar$ and $\omega$ equal to 1

First we need to initialize the parameters.
```{r}
  n = 5
  z = -100:100
```


We will have the compute the Hermite polynomial $H_n$ using the recurrence formula:
$$H_0 (z) = 1$$
$$H_1 (z) = 2z$$
$$\forall n \geqslant 1 , H_{n+1}(z) = 2zH_n(z) - 2nH_{n-1}(z)$$

```{r}
  H = matrix(nrow=length(z),ncol=n)
  H[,1] = 1
  H[,2] = 2*z
  for (i in 3:n) {
    H[,i] = 2*z*H[,i-1] - 2*(i-1)*H[,i-2]
  }
```

Computations of some terms in the Wave function
the variable t2 represents $(\frac {1}{\pi})^\frac {1}{4}$ and the zcarre represents the $z^2$ and exp represents $e^{-\frac{m\omega z^2}{2\hbar}}$
```{r}
  t2 = (1/pi)^0.25
  zcarre = z*z;
  exp = exp((-1*zcarre)/2)
```

Initialization of the $\Psi$ matrix
```{r}
  W = matrix(nrow=length(z), ncol=n)
```

Loop for each $n$ and compute the $\Psi_n$.

```{r}
  for (i in 1:n) {
    first_term = 1/sqrt((2^(i-1))*factorial(i-1))
    W[,i] = first_term * t2 * exp * H[,i]
  }
```

We show the data in the $\Psi$ matrix

```{r}
  W
```

It's almost impossible to get any sense out of this data so we will plot the matrix and visualize the data.

```{r}
  data = read.csv('../resources/out.csv')
  H = data[2:length(data)]
  z = data[1]
```

```{r}
  x = seq(-100, 100, 20)
  y = x^2
  matplot(z, H[,3], type="o", col="black", lwd = 3,xlim = c(-4,4), ylim = c(-1.5, 1.5))
  abline(h=0)
  abline(v=0)
  colors=c("#e6194B", "#3cb44b", "#4363d8", "orange", "brown");
  
  for (i in 2:ncol(H)) {
    lines(z, H[,i], type = "o", col="red", lwd = 3,ylim = c(-length(z), length(z)))
  }
  
  legend("topright", legend=c("W1", "W2", "W3", "W4", "W5"), col=colors, lty=1:2, cex=0.8)
```

